---
title: "Assignment 6- Covid-19 and bankruptcyies"
authur: "Andreas Knutsen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(dplyr)
library(lubridate)

bankrup_2019<- read_html("https://w2.brreg.no/kunngjoring/kombisok.jsp?datoFra=01.01.2019&datoTil=31.12.2019&id_region=0&id_niva1=51&id_niva2=56&id_bransje1=0")
bankrup_2020<- read_html("https://w2.brreg.no/kunngjoring/kombisok.jsp?datoFra=01.01.2020&datoTil=01.10.2020&id_region=0&id_niva1=51&id_niva2=56&id_bransje1=0")


```



## Dose Covid-19 impact the number of bankruptcies?

With the help of Brønnøysundregisteret we can find out how many bankruptcies there was in 2019 and what have happend in 2020, to see if there has been major changes.

### My codes.

This bit of the report will show you how I worked my way to find a answer to my question.

```{r}

#finding the data from 2019 and 2020 then combind them into one dataset.
bankrup_2019<- bankrup_2019%>%
  html_nodes(xpath = "//table")%>%
  html_nodes("table")%>%
  html_table(fill=TRUE)

bankrup_2020<- bankrup_2020%>%
  html_nodes(xpath = "//table")%>%
  html_nodes("table")%>%
  html_table(fill=TRUE)

bankrups<- bind_rows(bankrup_2019, bankrup_2020)

# selecting the columns we want to use, make a region column, just selecting the firms in X4.

bankrups<- bankrups%>%
  select(X2, X4, X6,X8)%>%
  mutate(region=ifelse(grepl("[^Dato][A-Za-z]", X6), X6,NA))%>%
  fill(region)%>%
  filter(nchar(X4)>=7)%>%
  filter(X8=="Konkursåpning")


bankrups<- bankrups%>%
  mutate(bankrups$X6<-as.Date(bankrups$X6, format = "%d.%m.%Y"))

colnames(bankrups)[6] <- "Dato"

#seperate the date variable
bankrups<-bankrups %>%
  mutate(
    dates2=ymd(Dato),
    År=year(dates2),
    Måned=month(dates2),
    Dag=day(dates2)
  )


#select again what we want to use and rename columns

bankrups1<- bankrups%>%
  select(X2,X8, region,År, Måned)
colnames(bankrups1)<-c("Firma","Kunngjøringstype","Fylke","År","Måned")

#count how many that have been bankrupted each mounth, year and county
bankrups1<- bankrups1%>%
  group_by(Fylke,Måned, År)%>%
  count(Kunngjøringstype)

colnames(bankrups1)[5] <- "Konkurs"

#convert År to a factor
bankrups1$År<-as.factor(bankrups1$År)


```

## The Plot

Now we will have a look how the number of bankruptcies are looking between 2019 and 2020.

```{r,message = FALSE, warrning= FALSE}
#plot it
bankrups1%>%
  ggplot(aes(x=Måned, y=Konkurs, group=År))+
  geom_line(aes(color=År),size=1)+
  facet_wrap(~Fylke)+
  scale_x_continuous(breaks=c(1:12))+
  ylab("Antall konkurser")+
  ggtitle("Antall konkurser i fylkene fra 01.01.2019 til i dag")
```


As we can see, it lookes like there is less bankruptcies in 2020 then in 2019. 