---
title: "Assignment 7- reading CDC web page"
author: "Andreas Knutsen"
date: "28 10 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(dplyr)
library(ggthemes)


```


## Childbirths in USA 2017-2019
This report will use the data from CDC webpage from 2017 to 2019 and look at three different topics. 
The first topic are "The proportion of boys to girls". Then we are looking at "The average birth weight in grams by gender". The last topic are "The proportion of boys to girls by day of birth".

Before we can look at the data to answer this topics, we have to read all of the data we need.

```{r,echo=TRUE, message = FALSE, warning= FALSE}
#I download all the data from the web page, and made a new variable Year.
# I turned all of the datasets from text to csv file. I did this because the files was very large, and the kniting took much longer then needed. I have marked the lines with #, so you can see what I did.
#birth2017 <- read_fwf("Nat2017PublicUS.c20180516.r20180808.txt",
#                      fwf_positions(start = c(13,23,475,504),
#                                    end = c(14,23,475,507),
#                                    col_names = c("Months","DayOfWeek", "Gender","Weight")
 #                     ))

#birth2017<- birth2017%>%
 # mutate(Year="2017")


#birth2018 <- read_fwf("Nat2018PublicUS.c20190509.r20190717.txt",
 #                     fwf_positions(start = c(13,23,475,504),
  #                                  end = c(14,23,475,507),
  #                                  col_names = c("Months","DayOfWeek", "Gender","Weight")
   #                   ))
#birth2018<- birth2018%>%
 # mutate(Year="2018")

#birth2019 <- read_fwf("Nat2019PublicUS.c20200506.r20200915.txt",
 #                     fwf_positions(start = c(13,23,475,504),
  #                                  end = c(14,23,475,507),
   #                                 col_names = c("Months","DayOfWeek", "Gender","Weight")
    #                  ))
#birth2019<- birth2019%>%
 # mutate(Year="2019")


#write.csv(birth2017,"birth2017.csv")

#rite.csv(birth2018,"birth2018.csv")


#write.csv(birth2019,"birth2019.csv")


birth2017csv<-read_csv("birth2017.csv")
birth2018csv<-read_csv("birth2018.csv")
birth2019csv<-read_csv("birth2019.csv")
#Bind all of the data, make weight numeric and Year character.

births<- bind_rows(birth2017csv, birth2018csv, birth2019csv)

births$Weight<- as.numeric(births$Weight)

births$Year<- as.character(births$Year)

```


### The proportion of boys to girls
```{r}
#Make a new datasets for girls and boys, and see how many is born each year. 

boys<- births%>%
  filter(Gender=="M")

allboys<- boys%>%
  group_by(Year)%>%
  tally()%>%
  mutate(Gender="M")
  


girls<- births%>%
  filter(Gender=="F")

allgirls<- girls%>%
  group_by(Year)%>%
  tally()%>%
  mutate(Gender="F")

boysgirls<-bind_rows(allboys,allgirls)

boysgirls%>%
  ggplot(aes(x=Gender, y=n, fill=Gender))+
  geom_bar(stat = "identity")+
  theme_economist()+
  facet_grid(~Year)+
  ggtitle("The proportion of boys to girls")+
  ylab("Number of births")


```

As we can see, there is born more boys then girls in this periode.

### "The average birth weight in grams by gender"

```{r}

#Find the average weight for boys and girls.

mean_boys<- boys%>%
  group_by(Year)%>%
  mutate(Average=mean(Weight))

mean_girls<- girls%>%
  group_by(Year)%>%
  mutate(Average=mean(Weight))

# find the average for boys and girls group by year, and slice the dataset to get just one value.

#Boys

meanb_17<- mean_boys%>%
  filter(Year=="2017")%>%
  select(Average, Year, Gender)%>%
  slice(1)
  

meanb_18<- mean_boys%>%
  filter(Year=="2018")%>%
  select(Average, Year, Gender)%>%
  slice(1)

meanb_19<- mean_boys%>%
  filter(Year=="2019")%>%
  select(Average, Year, Gender)%>%
  slice(1)

#Girls

meang_17<- mean_girls%>%
  filter(Year=="2017")%>%
  select(Average, Year, Gender)%>%
  slice(1)

meang_18<- mean_girls%>%
  filter(Year=="2018")%>%
  select(Average, Year, Gender)%>%
  slice(1)

meang_19<- mean_girls%>%
  filter(Year=="2019")%>%
  select(Average, Year, Gender)%>%
  slice(1)

#combind and ploting them

mean_weight<- bind_rows(meanb_17,meanb_18,meanb_19, meang_17, meang_18,meang_19)


mean_weight%>%
  ggplot(aes(x=Gender, y=Average, fill=Gender))+
  geom_bar(stat = "identity")+
  theme_economist()+
  facet_grid(~Year)+
  ggtitle("The average birth weight in grams by gender")


```

As we can see, the boys have higher average weight then girls.

### "The proportion of boys to girls by day of birth"

```{r}

boys17_days<- birth2017csv%>%
  filter(Gender=="M")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="M")%>%
  mutate(Year="2017")

boys18_days<- birth2018csv%>%
  filter(Gender=="M")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="M")%>%
  mutate(Year="2018")

boys19_days<- birth2019csv%>%
  filter(Gender=="M")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="M")%>%
  mutate(Year="2019")

girls17_days<- birth2017csv%>%
  filter(Gender=="F")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="F")%>%
  mutate(Year="2017")

girls18_days<- birth2018csv%>%
  filter(Gender=="F")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="F")%>%
  mutate(Year="2018")

girls19_days<- birth2019csv%>%
  filter(Gender=="F")%>%
  group_by(DayOfWeek)%>%
  tally()%>%
  mutate(Gender="F")%>%
  mutate(Year="2019")

#Combind and ploting them

birthdays<-bind_rows(boys17_days,boys18_days, boys19_days, girls17_days, girls18_days, girls19_days)

birthdays%>%
  ggplot(aes(x=Gender, y=n, fill=Year))+
  geom_bar(stat = "identity")+
  theme_economist()+
  facet_grid(~DayOfWeek)+
  scale_y_continuous(limits = c(0, 1000000))+
  ggtitle("The proportion of boys to girls by day of birth ")+
  ylab("Number of births")


```

From this data sunday is 1, monday 2 etc. As we can see, most are born in weekdays, monday to friday.

